Meta-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates

### 2025-02-13 paper/1_Initial/manuscript.qmd
updated manuscript

The IAS was also correlated with primal world beliefs, as measured by the Primal Inventory [PI-18; @clifton2021brief], which assesses beliefs about the world being alive, good, safe, and enticing. 
Items that evaluate neutral beliefs about the hierarchical order of importance in the world (i.e., hierarchical), as well as beliefs about the comprehensibility of most things and situations (i.e., understandable), and the belief that the world is characterized by flux (i.e., changing) were added as well.

@@

@@

>>>>

The IAS was also correlated with primal world beliefs, as measured by the Primal Inventory [PI-18; @clifton2021brief], which assesses beliefs about the world being alive, good, safe, and enticing. 
Items that evaluate neutral beliefs about the hierarchical order of importance in the world (i.e., hierarchical), as well as beliefs about the comprehensibility of most things and situations (i.e., understandable), and the belief that the world is characterized by flux (i.e., changing) were added as well.
The scale contains 18 items ranging from 5 (Strongly agree) to 0 (strongly disagree).

@@

The Generic Conspiracist Beliefs Scale [GCBS; @brotherton2013measuring] was used to assess five facets of conspiracy beliefs: Extraterrestrial, Global Conspiracies, Government Malfeasance, Information Control, and Personal Wellbeing. 
The scale comprises 15 items rated on a 5-point Likert scale, ranging from definitely not true (1) to definitely true (5).

Overall, the GCBS showed a weak but positive correlation with all facets of the IAS, with the strongest correlation observed between Global Conspiracies and Hungry/Thirsty (r = 0.140). 
Negative correlations were found within the Global Conspiracies, Extraterrestrial, and Information Control facets, though these were small, ranging from r = -0.0101 to r = -0.0167.

@@

Lastly, the Lying Profile Questionnaire [LIE; @makowski2023structure] a 16 item visual analog scale was used to assess 4 dispositional lying simensions: ability; negativity, contextuality, and frequency.

Overall, most lying profile dimensions show weak correlations with IAS pairings. 
Ability exhibits primarily weak positive correlations, with the strongest observed for Wind/Burp (r = 0.082).
In contrast, Frequency tends to show weak negative correlations, ranging from Wind/Burp (r = -0.062) to Muscles/Pain (r = -0.088).
Contextuality displays mixed correlations, with Itch/Bruise showing the strongest negative association (r = -0.127), while Urinate/Defecate has a small positive correlation (r = 0.045).
Finally, Negativity is consistently positively correlated with all pairings, with the strongest relationship found for Hungry/Thirsty (r = 0.090).

### 2025-02-13 paper/1_Initial/manuscript.tex
updated manuscript

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

@@

\par{     \begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colback=white, breakable, opacityback=0, leftrule=.75mm, arc=.35mm, bottomrule=.15mm, toprule=.15mm, left=2mm, colframe=quarto-callout-note-color-frame]

@@

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\begin{center}
\includegraphics[width=0.2\textwidth,height=\textheight]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
\end{center}  Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Ana
Neves:   Project administration, Data curation, Formal
Analysis, Investigation, Visualization, Writing -- original
draft, Writing -- review \& editing}
\par{Correspondence concerning this article should be addressed to }
}

@@

\includegraphics[width=6.4in,height=\textheight]{figures/MeasuringInteroception.jpg}

@@

\includegraphics[width=6.67in,height=\textheight]{figures/Figure1.png}

@@

The studies within our sample used the Body Perception Questionnaire
short-form (BPQ-SF) and the very-short form (BPQ-VSF) to assess
interoception (\citeproc{ref-cabrera2018assessing}{Cabrera et al.,
2018}). The BPQ-SF comprised of 46 items on a 5-point Likert scale
assessing body awareness (26 items) and autonomic reactivity (21 items).
The BPQ-VSF comprises of 12 items from the body awareness subscale of
the BPQ-SF. In this study, all scores assessing these two dimensions
were grouped together, hence no disitintion is made between awareness
measured with the BPQ-SF and the BPQ-VSF.

@@

The PHQ-15 also showed its strongest correlation with the Hungry/Thirsty
pair (r = -0.241) but was, on average, weakly and negatively correlated
with all other IAS pairings.

Lastly, the MSI-BPD also showed its strongest negative correlation with
the Hungry/Thirsty pair (r = -0.140) and was negatively correlated with
all other pairings, except for Cough/Sneeze, which showed a slight
positive correlation (r = 0.0219).

@@

{\centering \includegraphics[width=9.48958in,height=\textheight]{figures/clipboard-1619164537.png}

@@

The relationship between interoception and lying profiles was also weak.
This contrasts with previous research suggesting associations between
interoception and deception (\citeproc{ref-makowski2023a}{Makowski et
al., 2023}), warranting further investigation.

@@

Interestingly, world beliefs demonstrated significant positive
correlations with interoception. While this relationship has not been
previously documented, other forms of belief, such as political
ideology, have been linked to interoception
(\citeproc{ref-ruisch2022a}{Ruisch et al., 2022b}). Further research is
needed to determine whether world beliefs, which shape our perception of
reality (\citeproc{ref-clifton2020}{Clifton, 2020}), are meaningfully
connected to interoception.

@@

@@

@@

@@

@@

@@

>>>>

\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

@@

\par{     \begin{tcolorbox}[enhanced jigsaw, colback=white, colframe=quarto-callout-note-color-frame, leftrule=.75mm, opacityback=0, arc=.35mm, toprule=.15mm, breakable, rightrule=.15mm, left=2mm, bottomrule=.15mm]

@@

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\begin{center}
\includegraphics[width=0.2\linewidth,height=\textheight,keepaspectratio]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
\end{center}  Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Ana
Neves:   Project administration, Data curation, Formal
Analysis, Investigation, Visualization, Writing -- original
draft, Writing -- review \& editing}
\par{Correspondence concerning this article should be addressed to }
}

@@

\includegraphics[width=6.4in,height=\textheight,keepaspectratio]{figures/MeasuringInteroception.jpg}

@@

\includegraphics[width=6.67in,height=\textheight,keepaspectratio]{figures/Figure1.png}

@@

The studies within our sample used the Body Perception Questionnaire
short-form (BPQ-SF) and the very-short form (BPQ-VSF) to assess
interoception (\citeproc{ref-cabrera2018assessing}{Cabrera et al.,
2018}). The BPQ-SF comprised of 46 items on a 5-point Likert scale
assessing body awareness (26 items) and autonomic reactivity (21 items).
The BPQ-VSF comprises of 12 items from the body awareness subscale of
the BPQ-SF. In this study, all scores assessing these two dimensions
were grouped together, hence no disitintion is made between awareness
measured with the BPQ-SF and the BPQ-VSF, or eith scores obtained only
using the awareness subscale.

@@

The PHQ-15 is a 15-item questionnaire that assesses somatic symptoms on
a 3-point scale (e.g., back pain). It exhibited its strongest
correlation with the Hungry/Thirsty pair (r = -0.241) and, on average,
showed weak negative correlations with all other IAS pairings.

Lastly, the MSI-BPD is a 10-item questionnaire used to assess
personality disorder, where items are rated on a dichotomous scale of 1
(present) and 0 (absent). The MSI-BPD also showed its strongest negative
correlation with the Hungry/Thirsty pair (r = -0.140) and was negatively
correlated with all other pairings, except for Cough/Sneeze, which
showed a slight positive correlation (r = 0.0219).

\paragraph{Correlations with psychopathology
measures.}\label{correlations-with-psychopathology-measures}

Maladaptive personality traits were assessed using the Personality
Inventory for DSM-5 Short Form {[}PID-5-SF; Thimm et al.
(\citeproc{ref-thimm2016personality}{2016}){]}, which measures five
domains: disinhibition, antagonism, detachment, negative affect, and
psychoticism. The scale consists of 25 items rated on a 4-point Likert
scale, ranging from 0 (very false or often false) to 3 (very true or
often true).

On average, all maladaptive personality traits assessed by the PID-5-SF
were weakly and negatively correlated with IAS pairings. The strongest
correlation was observed between the psychoticism dimension and the
Muscle/Pain pairing (r = -0.173).

Schizotypy was assessed using the Schizotypal Personality Questionnaire
-- Brief Revised Updated (SPQ-BRU; Davidson et al.
(\citeproc{ref-davidson2016schizotypal}{2016})), which consists of 32
items rated on a 5-point Likert scale ranging from strongly agree to
strongly disagree. This questionnaire evaluates four primary dimensions:
cognitive-perceptual (positive), interpersonal (negative), disorganized,
and social anxiety. These dimensions are further divided into nine
secondary factors: constricted affect, eccentricity, magical thinking,
lack of close friends, odd speech, referential thinking, social anxiety,
suspiciousness, and unusual perceptions.

On average, all nine factors were weakly and negatively correlated with
IAS pairings, with correlations ranging from r = -0.170 (between lack of
close friends and Muscles/Pain) to r = 0.102 (between magical thinking
and Itch/Bruise).

The short version of the Autism-Spectrum Quotient (ASQ-Short; Hoekstra
et al., 2011) was used to assess five autistic traits: social skills,
adherence to routines, cognitive flexibility (switching), imagination,
and patterns/numbers. The questionnaire consists of 28 items rated on a
4-point Likert scale, ranging from 1 (definitely agree) to 4 (definitely
disagree).

Overall, all pairings were weakly and negatively correlated with the ASQ
dimensions, except for the Itch/Bruise and Heart/Breathing pairings,
which showed weak positive correlations with the patterns/numbers trait
(r = 0.184 and r = 0.038, respectively). The strongest correlation was
observed between the imagination trait and the Wind/Burp pairing (r =
-0.218).

\paragraph{Correlations with personality
measures.}\label{correlations-with-personality-measures}

The Big Five Inventory-Short Form {[}BFI-S; Lang et al.
(\citeproc{ref-lang2011short}{2011}){]} and the Mini International
Personality Item Pool {[}Mini-IPIP6; Sibley et al.
(\citeproc{ref-sibley2011mini}{2011}){]} were used to assess general
personality traits. The BFI-S consists of 15 items rated on a 7-point
Likert scale, ranging from 1 (strongly disagree) to 7 (strongly agree),
measuring five personality factors: Openness, Conscientiousness,
Extraversion, Agreeableness, and Neuroticism. The Mini-IPIP6 assesses
six personality traits---Extraversion, Agreeableness, Conscientiousness,
Neuroticism, Openness, and Honesty-Humility---using 24 items. While this
questionnaire is typically scored on a 7-point Likert scale from 1 (very
inaccurate) to 7 (very accurate), an analogous scale was used in the
respective sample. Lastly, the Neuroticism subscale of the NEO
Five-Factor Inventory {[}Neo-FFI; Costa and McCrae
(\citeproc{ref-costa1992normal}{1992}){]} was used to assess
Neuroticism, consisting of 12 items rated on a 5-point Likert scale from
1 (strongly disagree) to 5 (strongly agree).

To assess correlations with the IAS pairings, scores were grouped across
personality dimensions due to the overlap among these traits (with the
exception of Honesty-Humility). On average, IAS pairings were positively
associated with most personality dimensions, though these correlations
were generally weak. The strongest correlation was observed between
Conscientiousness and the Hungry/Thirsty pairing (r = 0.164). Both
Honesty-Humility and Neuroticism were weakly and negatively correlated
with the IAS pairings, with the highest correlations observed for the
Hungry/Thirsty pairing (r = -0.217 and r = -0.200, respectively).

\paragraph{Correlations with other
measures.}\label{correlations-with-other-measures}

The IAS was also correlated with primal world beliefs, as measured by
the Primal Inventory {[}PI-18; J. D. Clifton and Yaden
(\citeproc{ref-clifton2021brief}{2021}){]}, which assesses beliefs about
the world being alive, good, safe, and enticing. Items that evaluate
neutral beliefs about the hierarchical order of importance in the world
(i.e., hierarchical), as well as beliefs about the comprehensibility of
most things and situations (i.e., understandable), and the belief that
the world is characterized by flux (i.e., changing) were added as well.
The scale contains 18 items ranging from 5 (Strongly agree) to 0
(strongly disagree).

Overall, most primal beliefs show weak positive correlations with all
pairings of the IAS. The strongest correlation is between the
hierarchical belief and the Hungry/Thirsty pairing (r = 0.181). Some
beliefs, however, exhibit negative correlations with certain pairings.
These negative correlations range from r = -0.0940 between the changing
belief and the Hungry/Thirsty pairing, to r = -0.00490 between the
Enticing belief and the Itch/Bruise pairing.

The Generic Conspiracist Beliefs Scale {[}GCBS; Brotherton et al.
(\citeproc{ref-brotherton2013measuring}{2013}){]} was used to assess
five facets of conspiracy beliefs: Extraterrestrial, Global
Conspiracies, Government Malfeasance, Information Control, and Personal
Wellbeing. The scale comprises 15 items rated on a 5-point Likert scale,
ranging from definitely not true (1) to definitely true (5).

Overall, the GCBS showed a weak but positive correlation with all facets
of the IAS, with the strongest correlation observed between Global
Conspiracies and Hungry/Thirsty (r = 0.140). Negative correlations were
found within the Global Conspiracies, Extraterrestrial, and Information
Control facets, though these were small, ranging from r = -0.0101 to r =
-0.0167.

Lastly, the Lying Profile Questionnaire {[}LIE; Makowski, Pham, et al.
(\citeproc{ref-makowski2023structure}{2023}){]} a 16 item visual analog
scale was used to assess 4 dispositional lying simensions: ability;
negativity, contextuality, and frequency.

Overall, most lying profile dimensions show weak correlations with IAS
pairings. Ability exhibits primarily weak positive correlations, with
the strongest observed for Wind/Burp (r = 0.082). In contrast, Frequency
tends to show weak negative correlations, ranging from Wind/Burp (r =
-0.062) to Muscles/Pain (r = -0.088). Contextuality displays mixed
correlations, with Itch/Bruise showing the strongest negative
association (r = -0.127), while Urinate/Defecate has a small positive
correlation (r = 0.045). Finally, Negativity is consistently positively
correlated with all pairings, with the strongest relationship found for
Hungry/Thirsty (r = 0.090).

@@

{\centering \includegraphics[width=9.48958in,height=\textheight,keepaspectratio]{figures/clipboard-1619164537.png}

@@

The relationship between interoception and lying profiles was also weak.
This contrasts with previous research suggesting associations between
interoception and deception (\citeproc{ref-makowski2023a}{Makowski, Lau,
et al., 2023}), warranting further investigation.

@@

Interestingly, world beliefs demonstrated significant positive
correlations with interoception. While this relationship has not been
previously documented, other forms of belief, such as political
ideology, have been linked to interoception
(\citeproc{ref-ruisch2022a}{Ruisch et al., 2022b}). Further research is
needed to determine whether world beliefs, which shape our perception of
reality (\citeproc{ref-clifton2020}{J. D. W. Clifton, 2020}), are
meaningfully connected to interoception.

@@

\bibitem[\citeproctext]{ref-brotherton2013measuring}
Brotherton, R., French, C. C., \& Pickering, A. D. (2013). Measuring
belief in conspiracy theories: The generic conspiracist beliefs scale.
\emph{Frontiers in Psychology}, \emph{4}, 279.

@@

\bibitem[\citeproctext]{ref-clifton2021brief}
Clifton, J. D., \& Yaden, D. B. (2021). Brief measures of the four
highest-order primal world beliefs. \emph{Psychological Assessment},
\emph{33}(12), 1267.

\bibitem[\citeproctext]{ref-costa1992normal}
Costa, P. T., \& McCrae, R. R. (1992). Normal personality assessment in
clinical practice: The NEO personality inventory. \emph{Psychological
Assessment}, \emph{4}(1), 5.

\bibitem[\citeproctext]{ref-davidson2016schizotypal}
Davidson, C. A., Hoffman, L., \& Spaulding, W. D. (2016). Schizotypal
personality questionnaire--brief revised (updated): An update of norms,
factor structure, and item content in a large non-clinical young adult
sample. \emph{Psychiatry Research}, \emph{238}, 345--355.

@@

\bibitem[\citeproctext]{ref-lang2011short}
Lang, F. R., John, D., Lüdtke, O., Schupp, J., \& Wagner, G. G. (2011).
Short assessment of the big five: Robust across survey methods except
telephone interviewing. \emph{Behavior Research Methods}, \emph{43},
548--567.

@@

\bibitem[\citeproctext]{ref-makowski2023structure}
Makowski, D., Pham, T., Lau, Z. J., Raine, A., \& Chen, S. A. (2023).
The structure of deception: Validation of the lying profile
questionnaire. \emph{Current Psychology}, \emph{42}(5), 4001--4016.

@@

\bibitem[\citeproctext]{ref-sibley2011mini}
Sibley, C. G., Luyten, N., Purnomo, M., Mobberley, A., Wootton, L. W.,
Hammond, M. D., Sengupta, N., Perry, R., West-Newman, T., Wilson, M. S.,
et al. (2011). The mini-IPIP6: Validation and extension of a short
measure of the big-six factors of personality in new zealand. \emph{New
Zealand Journal of Psychology}, \emph{40}(3).

@@

\bibitem[\citeproctext]{ref-thimm2016personality}
Thimm, J. C., Jordan, S., \& Bach, B. (2016). The personality inventory
for DSM-5 short form (PID-5-SF): Psychometric properties and association
with big five traits and pathological beliefs in a norwegian population.
\emph{BMC Psychology}, \emph{4}, 1--11.
